---
title: "MA [46]15: Lab activity 1"
author: "Hongyi Yu"
output: html_document
---

# Introduction

In this lab activity we do a preliminary analysis of global social and economic indices. We use the *Social Progress Index* (SPI) from 2017 and the *Gross Domestic Product* (GDP) per capita from 2018. We downloaded the [social](https://www.socialprogressindex.com/assets/downloads/2017-Results.xlsx.zip) and [economic](http://api.worldbank.org/v2/en/indicator/NY.GDP.PCAP.PP.CD?downloadformat=excel) index data and put them at the `data` folder. We can now load them, along with the `tidyverse` package: 

```{r init, warning = FALSE}
library(tidyverse)
spi <- readxl::read_xlsx("data/2017 Results.xlsx") %>%
  mutate(spi = `Social Progress Index`)
ief <- readxl::read_xls("data/index2018_data.xls") %>%
  mutate(score = as.numeric(`2018 Score`), # Index of Economic Freedom (IEF)
         gdp = as.numeric(`GDP per Capita (PPP)`),
         Region = factor(Region))
```

We can now join the two datasets using `Country`:

```{r join, message = FALSE}
(ief.spi <- left_join(select(spi, Country, spi),
                     select(ief, Country, Region, score, gdp)))
```


# Visualization

Our first visualization is SPI against GDP:

```{r spi-gdp, warning = FALSE}
ggplot(ief.spi, aes(gdp, spi)) + geom_point(aes(color = Region)) +
  labs(x = "2018 GDP per capita (PPP)", y = "2017 Social progress index") +
  geom_smooth()
```

Our next visualization plots SPI against IEF:

```{r spi-ief, warning = FALSE}
ggplot(ief.spi, aes(score, spi)) + geom_point(aes(color = Region)) +
  labs(x = "2018 Index of economic freedom", y = "2017 Social progress index") +
  geom_smooth()
```

# Discussion

When GDP per capita is smaller than 50000, an increase in GDP per capita is associated with an increase in social progress index. When GDP per capita is greater than 50000, an increase in GDP per capita is associated with a decrease in social progress index.

When index of economic freedom is greater than 45, an increase in index of economic freedom is associated with an increase in social progress index. When index of eonomic freedom is smaller than 45, an increase in index of economic freedom is associated with a decrease in social progress index.