---
title: "MA [46]15: Lab Activity 5"
author: "Group: tidyverse"
output: github_document
---

**`[==[`**
In this lab activity you'll be working as a group. The data to be analyzed is
`data/world_pop.csv`, and contains information on world's population by
country. Your task is to verify [Benford's
law](https://en.wikipedia.org/wiki/Benford%27s_law), an empirical law saying
that the frequency of occurrence of first digits has a specific pattern, using
this dataset. To this end:

1. In a **Introduction** section, put a simple description of the dataset
(read the comment line in the CSV file for a reference) and about Benford's
law. Load the CSV into a table called `world_pop`.
2. In a **Data Analysis** section, compute for each `un_continental_region`,
the proportion of times `1` was a first digit among all countries in 2017, the
proportion of times `2` was a first digit, and so on. Store your results in a
table called `world_prop`. Using `world_prop`, make a line plot of the
frequencies (y-axis) per digit (x-axis), coloring the lines by
`un_continental_region`.
3. Compute the expected frequencies according to Benford's law and compare
them to the frequencies you obtained in the data analysis. Write a short
paragraph about your conclusions in a **Discussion** section.
The main goal here is **to work as a group**. There will be a single
repository for the whole group, so in the end you have to push a single
report. I strongly recommend that you elect a project manager that will merge
conflicts when group members push conflicting commits. Don't forget to use **branches**!
**`]==]`**

# Introduction
This is a list of countries and other inhabited territories of the world by total population, based on estimates published by the United Nations in the 2022 revision of World Population Prospects.
Benford's law states that in many real-life sets of numerical data, the leading digit is likely to be small.Benford's law also makes predictions about the distribution of second digits, third digits, digit combinations, and so on. We will verify Benford's law using the dataset.

```{r intro}
library(tidyverse)
world_pop <- read.csv("data/world_pop.csv", skip=1, header=TRUE)
```

# Data Analysis
```{r}
(world_prop <- world_pop %>%
  mutate(char_22 = as.character(population_2022), char_23 = as.character(population_2023), 
         first_23 = substr(population_2023, 1, 1)) %>% 
  group_by(un_continental_region, first_23) %>% 
  summarize(count = n()) %>% 
  mutate(prop = count / sum(count)))

world_prop %>% 
  ggplot(aes(first_23, count, group=un_continental_region)) + geom_line(aes(color = un_continental_region))

```

# Discussion
```{r}
benford <- tribble(~d, ~P_d,
                   #--#-----#
                   1, 0.301,
                   2, 0.176,
                   3, 0.125,
                   4, 0.097,
                   5, 0.079,
                   6, 0.067,
                   7, 0.058,
                   8, 0.051,
                   9, 0.046)
```

```{r}
(e_counts <- world_prop %>% 
   mutate(region_total = sum(count)) %>% 
   ungroup() %>% 
   add_row(un_continental_region = "Africa", first_23 = "7", count = 0, prop = 0.0, region_total = 58) %>% 
   add_row(un_continental_region = "Oceania", first_23 = "6", count = 0, prop = 0.0, region_total = 23) %>% 
   add_row(un_continental_region = "Oceania", first_23 = "8", count = 0, prop = 0.0, region_total = 23) %>% 
   mutate(E_count = region_total * benford$P_d[as.numeric(first_23)]) %>% 
   arrange(un_continental_region, first_23))
```
```{r}
(dists <- e_counts %>% 
  mutate(dist_i = (count - E_count)^2) %>% 
  group_by(un_continental_region) %>% 
  summarize(euclid_dist = sqrt(sum(dist_i))) %>% 
  arrange(euclid_dist))
```
