---
title: "MA [46]15: Lab Activity 3"
author: "Hongyi Yu"
output: github_document
---

# Introduction

Bike sharing systems are new generation of traditional bike rentals where
whole process from membership, rental and return back has become automatic.
Through these systems, user is able to easily rent a bike from a particular
position and return back at another position. Currently, there are about over
500 bike-sharing programs around the world which is composed of over 500
thousands bicycles. Today, there exists great interest in these systems due to
their important role in traffic, environmental and health issues.

We analyze the bike sharing dataset provided by Capital Bikeshare and
available as a [UCI ML repository](http://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset). 
The data span bike sharing activity in the Washington DC metro area for two years, 2011 and 2012. We start by loading the data:

```{r intro, message = FALSE}
library(tidyverse)
bike <- read_csv("/Users/victoria/Desktop/MA415/lab3-yhy808/data/bike-sharing.csv")
```

The variable `dteday` means date. 
The variable `holiday` means whether day is holiday or not. 
The variable `workingday` means if day is neither weekend nor holiday is 1, otherwise is 0.
The variable `weathersit` means 
  - 1: Clear, Few clouds, Partly cloudy, Partly cloudy
  - 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
  - 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
  - 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
The variable `temp` means  Normalized temperature in Celsius. The values are derived via (t-t_min)/(t_max-t_min), t_min=-8, t_max=+39.


# Data analysis

```{r a1}
(bike_fahr <- bike %>% mutate(instant, temp_F = 1.8*(temp*47-8)+32))
sum(bike_fahr$cnt) == sum(bike_fahr$casual) + sum(bike_fahr$registered)
```
Since we get "TRUE", we have checked that the total number of bike rentals corresponds to the number of casual and registered rentals.



```{r a2}
bike %>% ggplot() + geom_point(aes(x= dteday, y = cnt, color = factor(weathersit)))
```

The number of bike rentals is lowest during the January in each year. It will increase from January to June each year and become the highest in the July. Then, the number of bike rentals will decrease from July to December each year. Generally, the number of bike rentals growths from 2011 to 2012. Also, when weather condition is good, which means the weather is clear and has few clouds, there will be more bike rentals. When weather condition is bad, which means rain and snow, the number of bike rentals will be small.


# Discussion

```{r a3}
bike_fahr %>% ggplot() + geom_point(aes(x= dteday, y = cnt, color = temp_F, shape = factor(workingday)))
```
The number of bike rentals increases when the temperature is warm, and the number of bike rentals decreases when the temperature is cold. There is no obvious pattern between the number of bike rentals and if the date is a holiday or weekend.
