---
title: "MA [46]15 Homework 3"
author: "Hongyi Yu"
output: github_document
---

## Question 1

My work is to perform a preliminary EDA of the **real state properties** in
Boston in 2023. To this end, I downloaded the "Property Assessment FY2023"
data file from the
[Analyze Boston](https://data.boston.gov/dataset/property-assessment) web page,
and loaded it into a new dataset called `property`.


```{r q1}
library(tidyverse)
property <- read_csv("/Users/victoria/Desktop/MA415/hw3-yhy808/data/fy2023-property-assessment-data.csv",
                     col_types = cols_only(
                       LUC = col_integer(),
                       LU = col_character(),
                       GROSS_TAX = col_double(),
                       GROSS_AREA = col_integer()))
problems(property)

```


## Question 2

There are many "unusual" values, including tax exempt properties. Next, I'm
filtering out these cases to get a new dataset, `property_rep`.


```{r q2}
(property_rep <- property %>% 
   group_by(LUC) %>% 
   filter(median(GROSS_TAX) < 5000000, n() >= 5, LU != "E", LU != "EA") %>%
   ungroup())

```

## Question 3

To summarize what I have in `property_rep`, here is a table with the
proportions of properties by land use (`LU`), ordered by proportion, and the
cumulative proportions.


```{r q3}
(property_rep %>% group_by(LU)) %>%
  summarize(count = n()) %>%
  mutate(prop = count/sum(count)) %>%
  arrange(desc(prop)) %>%
  mutate(cul_prop = cumsum(prop))
```
The "LU" with the highest count is "CD", which means residential condominium unit. It occupies 40.87% of the total properties. Residential-1-family is the second high "LU" in the count of property, which is much higher than residential-2-family and residential-3-family. This result is expected because it is reasonable that residential units occupy a great part of properties.


## Question 4

Finally, I summarize my findings with log gross tax boxplots for each land
use.

```{r q4}
property_rep %>% group_by(LU) %>% 
  mutate(count = n()) %>% 
  ggplot(aes(reorder(LU, GROSS_TAX, FUN = median), log(GROSS_TAX))) + 
  geom_boxplot(aes(fill = count))
```

