---
title: 'MA [46]15: Lab activity 2'
author:
- name: Hongyi Yu
  affiliation: AAA
output:
  html_document: default
  pdf_document: default
---


# Introduction

In this lab activity we are going to analyze the `babynames` dataset.

**`[==[`** After loading the `babynames` package below, read the help for the `babynames` dataset and provide a short sentence describing it. **`]==]`**

We start by loading packages, including the `babynames` package containing the data:

```{r init, warning = FALSE}
library(tidyverse)
library(babynames) # install it with `install.packages("babynames")`
babynames
```


# Summarizing and visualizing the data 

Here are the names that were most popular in the whole record for both boys and girls:

```{r jointrank1}
babynames %>% group_by(sex, name) %>% summarize(total = sum(n)) %>%
  group_by(sex) %>% mutate(rank = min_rank(desc(total))) %>% arrange(rank, sex)
```

For comparison, here are the popular names since 1990, separately for boys and girls:

**`[==[`** Edit the commands below to produce two tables, one for boys and other for girls, with the most popular names since 1990, ordered by number of births. **`]==]`**

```{r jointrank}
babynames %>% filter(sex == "F", year >= 1990) %>% group_by(sex, name) %>% summarize(total = sum(n)) %>% mutate(rank = min_rank(desc(total))) %>% arrange(rank)
babynames %>% filter(sex == "M", year >= 1990) %>% group_by(sex, name) %>% summarize(total = sum(n)) %>% mutate(rank = min_rank(desc(total))) %>% arrange(rank)

```

The following `births.year` dataset summarizes the total number of births per sex across years:

```{r births-year}
births.year <- babynames %>% group_by(year, sex) %>% summarize(count = sum(n)) %>%
  spread(key = sex, value = count)
```

Here is a plot showing how the number of births varies yearly. The ratio of girls to boys is also plotted in color.

**`[==[`** Modify the command in `births-year-plot` to plot the total number of births per year and color the points using the proportion of girls to boys. **`]==]`**

```{r births-year-plot}
births.year %>% ggplot() + geom_point(aes(x = year, y = F + M, color = F/M))
```


**`[==[`**
Open a new Discussion section and summarize your findings in this short analysis. In particular:

  * Comment on possible changes to popular names in the last 20 years when compared to the most popular names in the whole record;
  
  * Comment on how the yearly number of births and ratio of girls to boys varies.
**`]==]`**

# Discussion

The popular female names in the last 20 years are Emily, Ashley, and Jessica, which are not the most popular names in the whole record. The popular female names in the whole record are Mary, Elizabeth, and Patricia. The popular male names in the last 20 years are Michael, Jacob, and Matthew. Michael is also a popular name in the whole record, but Jacob and Matthew are not popular in the whole record. The popular male names in the whole record are James, John, and Robert. It shows that the popular names in the last 20 years are different from the popular names in the whole record.

The number of births grows slowly before 1900. The ratio of girls to boys increases before 1900, which means the number of girls is greater than boys. Then, the number of births increases dramatically from 1900 to 1920, and the ratio of girls to boys decreases. After 1920, the number of births decreases slightly and then increases rapidly until 1960, and the ratio of girls to boys is close to 1. After 1960, the number of births keeps fluctuating but generally goes up, and the ratio of girls to boys is close to 1, which means the numbers of girls and boys are almost the same.


  