---
title: "MA [46]15 Homework 7"
author: "Hongyi Yu"
output: github_document
---


## Question 1


```{r q1}
library(tidyverse)
library(nycflights13)
library(geosphere)

dist_geo <- function (lon_from, lat_from, lon_to, lat_to) {
  geosphere::distGeo(cbind(lon_from, lat_from), c(lon_to, lat_to))
}

jfk <- airports %>% 
  filter(faa == "JFK") %>% 
  select(lon = lon, lat = lat)

(flights_dist <- airports %>% 
  filter(faa %in% unique(flights$dest)) %>%
  mutate(dist=dist_geo(lon, lat, jfk$lon, jfk$lat)) %>% 
  inner_join(flights, by = c("faa" = "dest")))

ggplot(flights_dist, aes(x = log(distance), y = log(dist))) + geom_point()

```


## Question 2

```{r q2}
flights_mt <- flights_dist %>%
  filter(!is.na(dep_delay) & !is.na(arr_delay) & !is.na(distance) & !is.na(tailnum)) %>%
  mutate(madeup_time = dep_delay - arr_delay)

lm_madeup_time <- lm(madeup_time ~ log(dist), flights_mt)

ggplot(flights_mt, aes(x = log(distance), y = madeup_time, alpha = 0.3)) + 
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = FALSE)

lm(madeup_time ~ log(dist), flights_mt)
lax <- airports %>% filter(faa == "LAX") %>% select(lon, lat)
log_dist_lax <- log(dist_geo(jfk$lon, jfk$lat, lax$lon, lax$lat)) 
madeup_time_lax <- coef(lm_madeup_time)[1] + coef(lm_madeup_time)[2] * log_dist_lax
(arr_delay_lax <- 30 - madeup_time_lax)

```
The linear coefficient is 2.132, which verifies the assumption that the travel will have more made-up time for delay if it has a long distance. 
If you're traveling to Los Angeles international airport and your flight left JFK with a 30 minute delay, the predicted arrival delay is about 22 minutes.


## Question 3

```{r q3}
flights_load <- flights %>%
  group_by(origin, month, day) %>% 
  summarize(trips = n())

rep_flights <- flights %>%
  group_by(carrier) %>% filter(n_distinct(flight) > 100) %>%
  group_by(carrier, flight) %>% filter(n() > 50) %>% ungroup()

(flights_load <- left_join(rep_flights, flights_load, by = c("origin", "month", "day")) %>%
  mutate(ontime = if_else(arr_delay <= 15 & !is.na(dep_time), 1, 0)))
```

## Question 4

```{r q4}
model <- glm(ontime ~ trips + carrier-1, flights_load, family = binomial)
(coef<-coef(model))

coef_carriers<-coef(model)[2:length(coef(model))]
carrier_data <- data.frame(
  carrier = c("9E","AA", "B6", "DL", "EV", "MQ", "UA", "US", "WN"),
  coefficient = coef_carriers)
ggplot(carrier_data, aes(x=fct_reorder(carrier, coefficient), y=coefficient)) +
  geom_bar(stat='identity')

((log(1)-coef[5])/coef[1])
```
The coefficient of trip is -0.006, which is negative. It means that an increase in the number of trips in a given date will decrease the probability that the flight being on-time, which meets my expectation. We need around 546 trips for the predicted probability of the flight being on-time to be 50%.
